---
title: "orfr"
output: github_document
---

## Model-Based Calibration and Scoring for Oral Reading Fluency Assessment Data with R

`orfr` is an R package that allows model-based calibration and scoring for oral reading fluency (ORF) assessment data.

### Installation:
To install `orf` package, follow the steps below.

1. The `MultiGHQuad` package needs to be installed. As of May 31, 2022, `MultiGHQuad` package has been removed from CRAN, so download an archive package file from  https://cran.r-project.org/src/contrib/Archive/MultiGHQuad/ and manually install.

2. Install `remotes` package, by running the following R code.
``` {r eval = F}
if(!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
```

3. Install `orfr` by running the following R code.
``` {r eval = F}
remotes::install_github("kamataak/orfr")
```

### Basic Usage:
It is recommended that the data are prepared as a long-format data frame, where each row is data for a unique case, namely, a specific passage from a specific student in a specific testing occasion. The data set should minimally contain the following variables: (1) student ID, (2) grade level, (3) testing occasion ID, (4) passage ID, (5) the number of words in the passage, (6) the number of words correctly read for the passage, and (7) time that took to read the passage.  

Variable names and the order of the variables can be flexible in the data frame. When running functions in this package, variable names for required variables need to be specified. 

To demonstrate some basic usage of key functions in the package, the `orfr` package contains an example passage-level student data set `passage`.     

Load required packages, and load/view the example data set `passage`.  
```{r eval = F}
library(tidyverse)
library(orfr)
data(passage)
View(passage)
```

#### Passage Calibration

Calibrate the passages using the `mcem()` function. 
``` {r eval = F}
test_MCEM <- mcem(stu.data=passage,
                  studentid = "id.student",
                  passageid = "id.passage",
                  numwords.p = "numwords.pass",
                  wrc = "wrc",
                  time = "sec",
                  k.in = 5,
                  reps.in = 2,
                  est = "mcem")
test_MCEM
```

#### Estimating WCPM scores

To estimate WCPM scores, we can do in three steps. 

**Step 1:** Prepare the data using the `prep()` function, where required variables for the `wcpm()` function are prepared, including changing variable names and a generation of the natural-logarithm of the time data. In addition, we can use another utility function `get.cases()` to generate a list of unique cases with the output of the `prep()` function. The output from prep() function has two lists: data.long is a long format of student response data, data.wide is a wide format of the natural-logarithm of the time data.   

```{r eval = F}
data <- prep(data = passage,
             vars=c("id.student","occasion","grade","id.passage","numwords.pass","wrc","sec"))
```

Generate a list of unique cases:
```{r eval = F}
get.cases(data$data.long)
```

**Step 2:** Specify a list of cases for which WCPM scores are estimated. It has to be a one-variable data frame with a variable name `cases`.  

The format of case values should be: `studentid_season`.  

```{r eval = F}
sample.cases <- data.frame(cases = c("2056_fall", "2056_winter", "2056_spring"))
```

If cases are not specified, the `wcpm()` function will estimate WCPM scores for all cases in the data in the next step.  

**Step 3:** Run the `wcpm()` function to estimate WCPM scores for selected cases. Note that we pass the output object `test_MCEM` from the passage calibration step, as well as the manipulated data `datalong` from Step 1. If the `cases =` argument is omitted, WCPM scores will be estimated for all cases in the data.

```{r eval = F}
test_WCPMEAP <- wcpm(calib.data=test_MCEM, 
                     stu.data = data$data.long,
                     cases = sample.cases, 
                     est = "map", 
                     se = "analytic")
summary(test_WCPM_MAP)
```

Also, we can specify a set of passages to scale the WCPM scores. If WCPM scores are scaled with a set of passages that is different from the set of passages the student read, the set of passages is referred to as an **external passage set**.  

The use of an external passage set is particularly important to make the estimated WCPM scores to be comparable between students for cross-sectional data, as well as within students for longitudinal data.

```{r eval = F}
test_WCPMEAP_EXT <- wcpm(calib.data=test_MCEM, 
                     stu.data = data$data.long,
                     cases = sample.cases, 
                     external = c("22007","22013","22036","22043","22048","22079"),
                     est = "map", 
                     se = "analytic")
summary(test_WCPM_MAP_EXT)
```

**Alternatively, we can run the `wcpm()` function without Step 1** above, by entering the original data `passage` directly as follows. 
``` {r eval = F}
test_WCPMEAP <- wcpm(calib.data=test_MCEM, 
                     stu.data = passage,
                     studentid = "id.student",
                     passageid = "id.passage",
                     season = "occasion",
                     grade = "grade",
                     numwords.p = "numwords.pass",
                     wrc = "wrc",
                     time = "sec",
                     cases = sample.cases, 
                     external = c("22007","22013","22036","22043","22048","22079"),
                     est = "map", 
                     se = "analytic")
summary(test_WCP_MEAP_ALT)
```

The package also contains calibrated passage data `MCEM`, in which passages were calibrated with much larger sample.  

Please see the [package website](https://kamataak.github.io/orfr/) for more detailed usage of the package.

### Citation


### Copyright Statement
Copyright (C) 2022 The ORF Project Team

The orfr package is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
 any later version.

The orfr package is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.
You should have received a copy of the GNU General Public License
along with this package.  If not, see <http://www.gnu.org/licenses/>.
